---
title: "CopernicusMarine Glossary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CopernicusMarine Glossary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Copernicus Marine Service organises its data in both NetCDF and Zarr
(a cloud-native array format) and discovered through a
[STAC (SpatioTemporal Asset Catalog) catalogue](https://stacspec.org/).
In order to handle the data efficiently it can be useful to understand
the terminology used by the data portal (adopted in the package).
Fortunately, the package tackles most technical challenges for you,
so you don't have to know all technical aspects.

However, since jargon can vary among data portals, I've put together
a glossary of terms used by Copernicus Marine. It is based on the
excellent [tutorial](https://r-rse.github.io/emodnet-bio-r-geo-tutorials/tutorials/tutorial-04.html)
by [Anna Krystalli](https://github.com/annakrystalli). This
tutorial also highlights some contrasts with other data portals.
The table below lists some important definitions and their background.

```{r glossary, echo=FALSE}
library(CopernicusMarine)
glossary <- cms_glossary()
glossary[] <- lapply(glossary, \(x) {
  x <- gsub("\\[(.*?)\\]\\((.*?)\\)", "<a href='\\2'>\\1</a>", x)
  x[is.na(x)] <- "-"
  x
})

DT::datatable(
  glossary,
  filter  = "top",
  escape  = FALSE,
  options = list(
    pageLength = 5,
    rownames   = FALSE,
    autoWidth  = TRUE,
    dom        = "tpi",
    columnDefs = list(
      list(targets = 0, visible = FALSE)
    )
  )
)
```
